---
title: "Module 8"
author: "Jimmy Erkens"
date: "`r Sys.Date()`"
output: 
    html_document:
     toc: true
     number_sections: true
     toc_float: true
     code_folding: show
     theme: journal
---

# Libraries

Libraries or whatever...
```{r, library}
library(tidyverse)
library(manipulate)
library(tinytex)
```


# Manipulate

Was not familiar with this package, actual goals, wish my stats profs used it in undergrad.

```{r}
# outcomes <- c(1, 2, 3, 4, 5, 6)
# manipulate::manipulate(hist(sample(outcomes, n, replace = TRUE), breaks = c(0.5, 1.5, 2.5,
   # 3.5, 4.5, 5.5, 6.5), probability = TRUE, main = paste("Histogram of Outcomes of ",
   # n, " Die Rolls", sep = ""), xlab = "roll", ylab = "probability"), n = slider(0,
   #  10000, initial = 10, step = 10))
```

# Challenge 1

```{r}
set.seed(812) # my birthday LEO GANG GANG
nrolls <- 1000
roll <- function(n) {
    sample(1:6, n, replace = TRUE) # sample with replacement go off; changed x to n for sake of ease
}
two_dice <- roll(nrolls) + roll(nrolls) # what happens when we take the average of the dice rolls????
avg_dice <- (roll(nrolls) + roll(nrolls))/2
hist(two_dice, breaks = c(1.5:12.5), probability = TRUE, main = "Rolling Two Dice",
    xlab = "sum of rolls", ylab = "probability")
hist(avg_dice, breaks = seq(from = 0, to = 6, by = 0.5), probability = TRUE, main = "Rolling Two Dice",
    xlab = "avg of rolls", ylab = "probability")
```

# Probability

Are we ever gonna do Bayes Rule? For the sake of the other students oh my god I hope not. _We're doing Baye's Rule oml_ <333
\
\
We go through (sorry not sorry) basic probability rules, been there done that.

# Challenge 2

P(face card) = 12/52
\
\
P(a King) = 4/52
\
\
P(a spade) = 13/52
\
\
P(spade | face) = 3/12
\
\
P(King | face) = 4/12
\
\
P(red AND face) = P(red | face) * P(face) = 1/2 * 12/52 = 6/52
\
\
P(club OR face$^c$) = P(club) + P(not a face) - P(club and face) = 43/52 (just copied your math I've done my math-stat community service)

# Discrete Random Variables

Coin flips and rolling a die

```{r}
outcomes <- c("heads", "tails")
prob <- c(1/2, 1/2)
barplot(prob, ylim = c(0, 0.6), names.arg = outcomes, space = 0.1, xlab = "outcome",
    ylab = "Pr(X = outcome)", main = "Probability Mass Function")
# im so juvenile bc cumsum makes me giggle still
cumprob <- cumsum(prob)
barplot(cumprob, names.arg = outcomes, space = 0.1, xlab = "outcome", ylab = "Cumulative Pr(X)",
    main = "Cumulative Probability")
```

Beta distribution

```{r}
a <- 2
b <- 1
K <- 2
x <- seq(from = 0, to = 1, by = 0.025)
fx <- K * x^(a - 1) * (1 - x)^(b - 1)
lower_x <- seq(from = -0.25, to = 0, by = 0.025)  # add some values of x less than zero
upper_x <- seq(from = 1, to = 1.25, by = 0.025)  # add some values of x greater than one
lower_fx <- rep(0, 11)  # add fx=0 values to x<0
upper_fx <- rep(0, 11)  # add fx=0 values to x>1
x <- c(lower_x, x, upper_x)  # paste xs together
fx <- c(lower_fx, fx, upper_fx)  # paste fxs together
d <- as.data.frame(cbind(x, fx))
p <- ggplot(data = d, aes(x = x, y = fx)) + xlab("x") + ylab("f(x)") + geom_line()
p
```

Cumulative function

```{r}
x <- seq(from = 0, to = 1, by = 0.005)
prob <- 0.5 * x * K * x^(a - 1) * (1 - x)^(b - 1)
barplot(prob, names.arg = x, space = 0, main = "Cumulative Probability", xlab = "x",
    ylab = "Pr(X ≤ x)")
```

