---
title: "Module 6"
author: "Jimmy Erkens"
date: "`r Sys.Date()`"
output: 
    html_document:
     toc: true
     number_sections: true
     toc_float: true
     code_folding: show
     theme: journal
---

# Preliminaries

## Loading `tidyverse`
```{r, packages, warning = F}
library(curl)
library(tidyverse)
```
## Loading the data using `curl`

```{r, data_init, warning = F}
country_data <- read_csv(curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/Country-Data-2016.csv")) # I disagree with the syntax in the module, we don't need to make a new variable f
```

Hey! Heads up the subheader "For univariate data" isn't formatted correctly

# Some useful procedures for data analysis

## Challenge 1

```{r}
summary(country_data$population) # median population is 4.912 x 10^6 people
summary(country_data$area) # median area is 69,700 

# using tidyverse to make a new column for population density
# here we use a pipe operator and the mutate() function to make it work
country_data <- country_data %>% mutate(pop_density = population/area)
country_data[order(country_data$pop_density, decreasing = T), ][1:10] # most dense countries!
country_data[order(country_data$pop_density, decreasing = F), ][1:10] # least dense countries!
```

## Challenge 2

Fun with plotting! Sorry! `ggplot2` is better :)

```{r}
par(mfrow = c(2, 3))
boxplot(country_data$population)
boxplot(log(country_data$population)) # log standardizes non-normal data (generally!) this is a massive generalization!
boxplot(country_data$area)
boxplot(log(country_data$area)) 
barplot(country_data$population)
barplot(country_data$area)
```

## Challenge 3

```{r}
par(mfrow = c(1, 2))  # gives us two panels
attach(country_data) # so this is what we do for when we don't use ggplot()
hist(log(population), freq = FALSE, col = "red", main = "Plot 1", xlab = "log(population size)",
    ylab = "density", ylim = c(0, 0.2))
hist(log(area), freq = FALSE, col = "red", main = "Plot 2", xlab = "log(area)",
    ylab = "density", ylim = c(0, 0.2))

par(mfrow = c(1, 1))  # set up one panel and redraw the log(population) histogram
hist(log(population), freq = FALSE, col = "white", main = "My Plot with Mean and Density",
    xlab = "log(population size)", ylab = "density", ylim = c(0, 0.2))
abline(v = mean(log(population), na.rm = TRUE), col = "blue")
lines(density(log(population), na.rm = TRUE), col = "green")
detach(country_data) # note how attach() and detach() work when we work with `base` packages
```

## Challenge 4

```{r}
sort(table(country_data$govt_form), decreasing = TRUE) # repubic is the most common!
```

I'm more than a little disgusted at the sheer number of overseas territories and dependencies. Republic is the most common!


## Challenge 5

```{r, challenge_5}
kandc <- read_csv(curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/KamilarAndCooperData.csv"))
attach(kandc)
head(kandc)
summary(kandc)
boxplot(log(Body_mass_female_mean) ~ Family, kandc)
detach(kandc)

# plotting using ggplot! The better way to plot!
library(ggplot2)
p <- ggplot(data = kandc, aes(x = Family, y = log(Body_mass_female_mean)))  #define the variables
p <- p + geom_boxplot()  #graph them in a boxplot
p <- p + theme(axis.text.x = element_text(angle = 90))  #put x-axis names at 90deg
p <- p + ylab("log(Female Body Mass)")  #rename y-axis title
p  #show me the graph
```

## Challenge 6

```{r, scatterplot, warning = F}
attach(kandc)
par(mfrow = c(1, 2))
plot(x = Body_mass_female_mean, y = Brain_Size_Female_Mean)
plot(x = log(Body_mass_female_mean), y = log(Brain_Size_Female_Mean))
```

Using `ggplot2()`! I get why we're initializing but I don't necessarily agree with this syntax for practical use.

```{r, ggplot}
p <- ggplot(data = kandc, aes(x = log(Body_mass_female_mean), y = log(Brain_Size_Female_Mean),
    color = factor(Family)))  # first, we build a plot object and color points by Family
p <- p + xlab("log(Female Body Mass)") + ylab("log(Female Brain Size)")  # then we modify the axis labels
p <- p + geom_point()  # then we make a scatterplot
p <- p + theme(legend.position = "bottom", legend.title = element_blank())  # then we modify the legend
p  # and, finally, we plot the object
```

Faceting!

```{r, facet}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```

## Challenge 7

```{r, bivariate}
p <- ggplot(data = kandc, aes(x = log(Body_mass_female_mean), y = log(MaxLongevity_m)))
p <- p + geom_point()
p <- p + geom_smooth(method = "lm")
p
```


# Aggregating Statistics

```{r}
aggregate(kandc$Body_mass_female_mean ~ kandc$Family, FUN = "mean", na.rm = TRUE) 
# grouping by mean! let's see how dplyr makes it easier
aggregate(x = kandc["Body_mass_female_mean"], by = kandc["Family"], FUN = "mean", na.rm = TRUE)
# grouping by female mean, note the difference in function
```

Going through the `dplyr` functions!

```{r}
s <- filter(kandc, Family == "Hominidae" & Mass_Dimorphism > 2)
head(s)  # filtering a data frame for certain rows...


s <- arrange(kandc, Family, Genus, Body_mass_male_mean)  # rearranging a data frame... NOTE how this is arranged
head(s)

s <- rename(kandc, Female_Mass = Body_mass_female_mean)
head(s$Female_Mass)  # renaming columns...

s <- mutate(kandc, Binomial = paste(Genus, Species, sep = " ")) # I feel there's another way to concatenate like this
head(s$Binomial)  # and adding new columns...

s <- summarise(kandc, avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
    na.rm = TRUE)) # 
s

# using group_by()
byFamily <- group_by(kandc, Family)
byFamily

s <- summarise(byFamily, avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
    na.rm = TRUE))
s # how does summarise() work with groups!

```

# Piping

```{r}
s <-                                                             #to create dataframe "s"
  kandc %>%                                                          #take dataframe "d"
  group_by(Family) %>%                                           #Group it by Family
  summarise(avgF = mean(Body_mass_female_mean, na.rm=TRUE),      #And calculate mean male BM
            avgM = mean(Body_mass_male_mean, na.rm=TRUE))        #And mean female BM
s
```

## Challenge 8

```{r}
s <- kandc %>%
    mutate(Binomial = paste(Genus, Species, sep = " ")) %>% # I honestly hate strings
    select(Binomial, Body_mass_female_mean, Body_mass_male_mean, Mass_Dimorphism) %>% # we take these groups
    group_by(Binomial) %>% # and then attach them to Binomial
    summarise(avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
        na.rm = TRUE), avgBMD = mean(Mass_Dimorphism, na.rm = TRUE)) # so we then are able to have our summary in accordance w group
s %>% filter(Binomial == "Chlorocebus pygerythrus" | Binomial == "Lagothrix lagotricha") # there's a better way to do this but I'm lazy
```

Vervet monkeys are more sexually dimorphic, which makes sense! Woolly monkeys have a larger female body mass!

```{r}
kandc %>% filter(Family == "Cercopithecidae" | Family == "Atelidae") %>% 
  ggplot(aes(x = Body_mass_male_mean, fill = Family)) +
  geom_boxplot() +
  scale_fill_manual(values = c("purple", "green")) + # this is absolutely hideous I can't... 
  ggtitle("Boxplot of average male mass by Family")

kandc %>% filter(Family == "Cercopithecidae" | Family == "Atelidae") %>% 
  ggplot(aes(x = Body_mass_female_mean, fill = Family)) +
  geom_boxplot() +
  scale_fill_manual(values = c("purple", "green")) + # this is absolutely hideous I can't... 
  ggtitle("Boxplot of average female mass by Family")

kandc %>% filter(Family == "Cercopithecidae" | Family == "Atelidae") %>% group_by(Family) %>% 
  ggplot(aes(x = Mass_Dimorphism, fill = Family)) +
  geom_boxplot() +
  scale_fill_manual(values = c("purple", "green")) + # this is absolutely hideous I can't... 
  ggtitle("Boxplot of average sexual dimorphism by Family")
```

